---
title: "The impact of gender on earnings project proposal"
author: "Group 3: Lorenzo Wahaus, Sonia Duan, Jingrui Zhou"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE, echo=FALSE}
library(tidyverse)
library(openintro)
library(broom)
library(scales)
jobs_gender <- read_csv(file="https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")
```

# Section 1 - Introduction

### Intro

The data that we will be using for our project is the jobs_gender data from the tidytuesday repo. The full url for the data is: "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv", and we found the data at: "https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-03-05/jobs_gender.csv". The data looks at the difference in earnings between men and women across various jobs from 2013-2016. The question we want to answer is what impact gender has on earnings, and if it varies for different jobs or fields.

### Other research

Women have historically been marginalized when it comes to jobs. There is a lot of history of paying women less than men for the same job. With more women joining the workforce than before, and gender equality still an important issue.

According to the U.S. Bureau of Labor Statistics in 2019 57.4% of all women were in the labor force, compared to 69.2% of all men. This is following the rapid rise of women joining the labor force in the latter half of the 20th century. Following this rise of women working, their earnings as a proportion of men's earnings rose as well. But, now we want to see how much of a gap there is left to close.

### Details on the data

The data comes from surveys conducted by the Census Bureau and the Bureau of Labor Statistics, and is estimates based on the data they collected. The collected data is based on jobs in America and the earnings of the workers in those jobs.

### Code Book
The United States Census Bureau an the Bureau of Labor Statistics takes surveys of people, their job, and their earnings, among other factors, to see what work force the country has available. The data set that we are using is based off of the jobs data from 2013-2016. The data set contains 2088 data cases.
```{r codebook, echo=FALSE}
variable <- c("year", "occupation", "major_categroy", "minor_category", "total_workers", "workers_male", "workers_female", "percent_female", "total_earnings", "total_earnings_male", "total_earnings_female", "wage_percent_female")
NorC <- c("categorical", "categorical", "categorical", "categorical", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric")
description <- c("year of the data point", "job", "broad category of job", "specific category of job", "estimated total of full time workers in job", "estimated total of male full time workers", "estimated total of female full time workers", "percent of estimated workers that are female", "estimated median earnings of all full time workers", "estimated median earnings of all male full time workers", "estimated median earnings of all female full time workers", "percent of men's wages that women earn")

codebook <- data.frame(variable, NorC, description)

codebook
```

### Data Glimpse

```{r glimpse, echo=FALSE}
glimpse(jobs_gender)
```

# Section 2 - Data Analysis Plan

### Potential Data Relationships

-Could the percent of workers in a job that are women(percent_female) impact how much women make compared to men(wage_percent_of_male)?

-Does the earnings of women compared to men(wage_percent_of_male) change through the years(year)?

-Which occupations(occupation) have the biggest gender pay gap(wage_percent_of_male)?

-Does the total amount of workers(total_workers) in a job impact how much women make(total_earnings_female)?

### Challenges

One potential challenge we could face is that the data only covers 2013 to 2016, no more than four years. This means that we will not be able to look at earnings over a long period of time.

Another challenge is that the data only looks at the job and the earnings. It does not look at other potential impacts on earnings such as education level or jobs available. This means that any earnings differences we find could potentially be a result of unequal access to education or jobs and not necessarily gender.

### Preliminary Exploratory Data Analysis

##### The proportion of female labor force in all job category are shown below.

```{r, echo=FALSE, warning=FALSE}
jobs_gender %>% 
  group_by(major_category) %>% 
  mutate(female_labor = workers_female / total_workers) %>% 
  summarise(mean_female_labor = percent(mean(female_labor, na.rm = TRUE))) 


```
By having such a visualization shown composition of labor force in different *job category*, we are therefore aware that the distribution of labor(with different gender) are not equivalent in different job category and could thus have a **more holistic and un-biased** analysis in our project.


##### Do women make less than men?
```{r echo=FALSE, warning=FALSE}
ggplot(data=jobs_gender, mapping=aes(x=wage_percent_of_male)) +
  geom_histogram(binwidth = 2) +
  labs(title="Histogram of women's wage as percent of men's",
       x="Wage as percent of men's in the same job",
       subtitle = "data from tidytuesday, see bibliography")
```

Looking at a histogram of women's wage as a percent of men's wage, the histogram has a bell curve, but the middle of the curve is not near _100%_. The middle of the curve is around 85%-95%, meaning that women appear to be making less than men in the same job. So there is something for us to look into.

# Section 3 - Bibliography

Jobs gender data: https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv

Women in the labor force: a databook: https://www.bls.gov/opub/reports/womens-databook/2020/home.htm#:~:text=In%202019%2C%2057.4%20percent%20of,of%2060.0%20percent%20in%201999.

...

