---
title: "Lab Name"
author: "Group 3: Lorenzo Wahaus, Sonia Duan, Jingrui Zhou"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE, echo=FALSE}
library(tidyverse)
library(openintro)
library(tidymodels)
jobs <- read_csv(file="jobs.csv")
```

```{r load_and_change_data, include=FALSE}
jobs_gender <- read_csv(file="https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")

jobs_altered <- jobs_gender %>% mutate(major_category = case_when(
  major_category == "Computer, Engineering, and Science" ~ "Computing",
  major_category == "Education, Legal, Community Service, Arts, and Media" ~ "Arts",
  major_category == "Healthcare Practitioners and Technical" ~ "Healthcare",
  major_category == "Management, Business, and Financial" ~ "Finances",
  major_category == "Natural Resources, Construction, and Maintenance" ~ "Construction",
  major_category == "Production, Transportation, and Material Moving" ~ "Transportation",
  major_category == "Sales and Office" ~ "Sales",
  major_category == "Service" ~ "Service"))

jobs_altered <- jobs_altered %>% select(year, major_category, workers_female, percent_female, total_earnings, total_earnings_female, wage_percent_of_male)

write_csv(jobs_altered, "jobs.csv")
```

```{r split_data, echo=FALSE}
set.seed(539)
split_data <- initial_split(jobs, prop=0.5)
training_data <- training(split_data)
test_data <- testing(split_data)
```

# Section 1 - Introduction

## Background

## Our Data

The data that we will be using for our project is the jobs_gender data from the tidytuesday repo. The full url for the data is: "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv", and we found the data at: "https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-03-05/jobs_gender.csv". The data looks at the difference in earnings between men and women across various jobs from 2013-2016.

The data comes from surveys conducted by the Census Bureau and the Bureau of Labor Statistics, and is estimates based on the data they collected. The collected data is based on jobs in America and the earnings of the workers in those jobs.

The United States Census Bureau an the Bureau of Labor Statistics takes surveys of people, their job, and their earnings, among other factors, to see what work force the country has available. The data set that we are using is based off of the jobs data from 2013-2016. The data set contains 2088 data cases.

For our modified data set, we removed any variables that we did not use, and also re-coded the `major_category` variable. In the original data set it was a long string on multiple words, and we cut it down to one word each to help conserve space on our visauals using `major_category`.

### Code Book

Code Book of Variables Used

Variable  | Categorical or Numeric | description
------------- | ------------- | --------
year  | numeric  |  year the data point is from
major_category  | categorical  | Large category that best fits the job of the data point
workers_female  | numeric  | Estimated total of female workers in the job
percent_female  | numeric  | Estimated percent of total workers in the job that are female
total_earnings  | numeric  | Estimated median earning for all workers in the job
total_earnings_female  | numeric  | Estimated median earnings for female workers in the job
wage_percent_of_male  | numeric  | Wages of female workers as a percent of male workers

# Section 2 - Model Building

## Lorenzo Model

### Percent of Workers That are Women

```{r differential, include=FALSE, eval=FALSE}
#showing how I made the differential file included
set.seed(539)
boot_perc_df <- training_data %>%
  specify(wage_percent_of_male ~ percent_female) %>%
  generate(reps=10000, type="bootstrap") %>%
  calculate(stat="slope")

write_csv(boot_perc_df, "perc_fem_df.csv")
```

One of the first variables that I thought of using to predict `wage_percent_of_male` was `percent_female`. This is because the more women there are in a job, the larger of a voice they have among the workers, and the percent of workers in a job that are female was one of the ways to measure that. Looking at the 95% confidence interval for a bootstrap differential of the slope with 10000 repetitions of the original data, both the upper bound and lower bound are positive. This means that there is a positive relationship between `percent_female` and `wage_percent_of_male`, and thus `percent_female` will be used in constructing my model.

```{r confidence_interval, echo=FALSE, message=FALSE, show_col_types=FALSE}
perc_fem_df <- read_csv(file="perc_fem_df.csv")
perc_fem_df %>% get_confidence_interval(level=0.95)
```

### Year

I also know that based on historical trends, women's wages have historically been less than men, with a trend towards wage equality. Although our data only has four years, I wanted to know if it also reflected this trend and could give me another variable for my model. I ran into an issue when making the box plot, where it would not separate based on year, so I had to make a new temporary variable `year_category` for the box plot. Looking at the box plot, the median `wage_percent_of_male` appears to increase each year, with the exception of 2016 having a slightly lower median than 2015. But, the general trend is still there, so I will be using year as my second variable.

```{r year_boxplot, echo=FALSE, warning=FALSE}
training2 <- training_data %>% mutate(year_category = case_when(
  year == 2013 ~ "Two Thousand Thirteen",
  year == 2014 ~ "Two Thousand Fourteen",
  year == 2015 ~ "Two Thousand Fifteen",
  year == 2016 ~ "Two Thousand Sixteen"))

ggplot(data = training2, 
       mapping=aes(x=wage_percent_of_male, fill=year_category)) +
  geom_boxplot() +
  labs(title="Wage Percentage Compared to Year",
       x="wage percent of male",
       fill="year") +
  theme(
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()
  ) +
  scale_fill_viridis_d()
```

### Category of Job

For my third variable, I wanted to use `major_category`. This is because different jobs pay different wages, and thus different categories of jobs could be paying women more or less equally when compared to men. To test this I made a box plot looking at `wage_percent_of_male`, and split the boxes up based on `major_category`. Looking at the box plot proved my hypothesis correct, as the boxes for Transportation, Sales, and Finances were noticeably lower than the other boxes. Furthermore, computing jobs had a much higher 25th percentile compared to the others. Combined with the large variety of whiskers on the plots, I felt that `major_category` was a good variable to include in my model.

```{r major_cat, echo=FALSE, warning=FALSE}
ggplot(data = training_data, mapping=aes(x=wage_percent_of_male, fill=major_category)) +
  geom_boxplot() +
  labs(title="Wage Percentage Compared to Category",
       x="Wage percent of male",
       fill="Category")+
  theme(
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()
  )+
  scale_fill_viridis_d()
```

### Building the Lorenzo Model

Combining all of my variables into a single linear model gets the equation:

wage_percent_of_male-hat = -828 + (0.046 * percent_female) + (0.452 * year) + (2.09 * Computing) + (1.80 * Construction) + (-4.71 * Finances) + (1.88 * Healthcare) + (-2.20 * Sales) + (1.46 * Service) + (-3.39 * Transportation)

```{r Lorenzo_model, echo=FALSE}
Lorenzo_model <- lm(wage_percent_of_male ~ percent_female + year + major_category, data = training_data)

Lorenzo_model %>% tidy()
```
```{r Lorenzo_stats, include=FALSE}
glance(Lorenzo_model) %>% select(r.squared, adj.r.squared)

Lorenzo_pred <- predict(Lorenzo_model, newdata=test_data) %>%
  bind_cols(test_data %>% select(wage_percent_of_male)) %>%
  rename(pred = ...1)

rmse(Lorenzo_pred, truth=wage_percent_of_male, estimate=pred)
```

## Sonia Model

## Jingrui Model

## Models Comparison

# Section 3 - Results

...

