---
title: "Are Female Workers Payed Equally?"
author: "Group 3: Lorenzo Wahaus, Sonia Duan, Jingrui Zhou"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE, echo=FALSE}
library(tidyverse)
library(openintro)
library(tidymodels)
jobs <- read_csv(file="jobs.csv")
```

```{r load_and_change_data, include=FALSE}
jobs_gender <- read_csv(file="https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")

jobs_altered <- jobs_gender %>% mutate(major_category = case_when(
  major_category == "Computer, Engineering, and Science" ~ "Computing",
  major_category == "Education, Legal, Community Service, Arts, and Media" ~ "Arts",
  major_category == "Healthcare Practitioners and Technical" ~ "Healthcare",
  major_category == "Management, Business, and Financial" ~ "Finances",
  major_category == "Natural Resources, Construction, and Maintenance" ~ "Construction",
  major_category == "Production, Transportation, and Material Moving" ~ "Transportation",
  major_category == "Sales and Office" ~ "Sales",
  major_category == "Service" ~ "Service"))

jobs_altered <- jobs_altered %>% select(year, major_category, workers_female, percent_female, total_earnings, total_earnings_female, wage_percent_of_male)

write_csv(jobs_altered, "jobs.csv")
```

```{r split_data, echo=FALSE}
set.seed(539)
split_data <- initial_split(jobs, prop=0.5)
training_data <- training(split_data)
test_data <- testing(split_data)
```

# Introduction

Tracing back to the early 20th century, most women in the United States only stayed at home taking care of babies or housekeeping. Even those who worked outside the home were primarily unmarried. **Women were trapped in homes at that time**. However, people often understate their contributions to the economy beyond family caring. According to our background research, we found that despite of **the limited opportunities available for women**, their participation in the economy still continued to rise between the mid 20 century. 

Nowadays, with more and more female workers entering the workforce than ever before, **we are interested in the wage equality between males and females**. When women became more aware of their own circumstances in their careers as well as their families, they began to **pay more attention on their own rights and benefits**. We want to focus on the gender equality of income levels in the project.

So we located a data set that was created from the Census Bureau and Bureau of Labor Statistics surveys. The information gathered is based on American jobs and the earnings of persons who work in such jobs. Searching the background information of the original data set, we found that the United States Census Bureau and the Bureau of Labor Statistics takes surveys of people, their jobs and their earnings. The data set that we're working with is based on job data from 2013 to 2016, containing a sample of 2088 data cases.

The full url for the original data set is: "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv",  and we found the data at: "https://github.com/rfordatascience/tidytuesday/blob/master/data/2019/2019-03-05/jobs_gender.csv".

We then created a code book of the variables that we planned to use to build each of our models.

Variable  | Categorical or Numeric | description
------------- | ------------- | --------
year  | numeric  |  year the data point is from
major_category  | categorical  | Large category that best fits the job of the data point
workers_female  | numeric  | Estimated total of female workers in the job
percent_female  | numeric  | Estimated percent of total workers in the job that are female
total_earnings  | numeric  | Estimated median earning for all workers in the job
total_earnings_female  | numeric  | Estimated median earnings for female workers in the job
wage_percent_of_male  | numeric  | Wages of female workers as a percent of male workers

For our modified data set, we removed the variables that we didn't use in the model building. Also, we re-coded the `major_category` variable. In the original data set it was a long string of multiple words, we cut it down to one word each to help conserve space on our visuals that used `major_category` as a variable.

For our **response variable**, we chose `wage_percent_of_male`.

We used it to compare the female workers income levels with those of male workers. It depicts female workers' wages as a percentage of male workers' wages. We can simply determine whether the gender wage equality is changing by evaluating the relationship between `wage percent of male` and other relevant variables.

**We want to discuss about the states of the women in their career basing on their income levels in comparison to men**. As people who wish the living conditions of females to improve in the United States, we believe it's a meaningful topic to discuss about. Because you cannot have gender equality if one gender is being payed less for their work.

# Model Building

## Model proposed by Lorenzo

```{r differential, include=FALSE, eval=FALSE}
#showing how I made the differential file included
set.seed(539)
boot_perc_df <- training_data %>%
  specify(wage_percent_of_male ~ percent_female) %>%
  generate(reps=10000, type="bootstrap") %>%
  calculate(stat="slope")

write_csv(boot_perc_df, "perc_fem_df.csv")
```

One of the first variables that I thought of using to predict `wage_percent_of_male` was `percent_female`. This is because the more women there are in a job, the larger of a voice they have among the workers, and the percent of workers in a job that are female was one of the ways to measure that. Looking at the 95% confidence interval for a bootstrap differential of the slope with 10000 repetitions of the original data, both the upper bound and lower bound are positive. This means that there is a positive relationship between `percent_female` and `wage_percent_of_male`, and thus `percent_female` will be used in constructing my model.

```{r confidence_interval, echo=FALSE, message=FALSE, show_col_types=FALSE}
perc_fem_df <- read_csv(file="perc_fem_df.csv")
perc_fem_df %>% get_confidence_interval(level=0.95)
```

### Year

I also know that based on historical trends, women's wages have historically been less than men, with a trend towards wage equality. Although our data only has four years, I wanted to know if it also reflected this trend and could give me another variable for my model. I ran into an issue when making the box plot, where it would not separate based on year, so I had to make a new temporary variable `year_category` for the box plot. Looking at the box plot, the median `wage_percent_of_male` appears to increase each year, with the exception of 2016 having a slightly lower median than 2015. But, the general trend is still there, so I will be using year as my second variable.

```{r year_boxplot, echo=FALSE, warning=FALSE}
training2 <- training_data %>% mutate(year_category = case_when(
  year == 2013 ~ "Two Thousand Thirteen",
  year == 2014 ~ "Two Thousand Fourteen",
  year == 2015 ~ "Two Thousand Fifteen",
  year == 2016 ~ "Two Thousand Sixteen"))

ggplot(data = training2, 
       mapping=aes(x=wage_percent_of_male, fill=year_category)) +
  geom_boxplot() +
  labs(title="Wage Percentage Compared to Year",
       x="wage percent of male",
       fill="year") +
  theme(
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()
  ) +
  scale_fill_viridis_d()
```

### Category of Job

For my third variable, I wanted to use `major_category`. This is because different jobs pay different wages, and thus different categories of jobs could be paying women more or less equally when compared to men. To test this I made a box plot looking at `wage_percent_of_male`, and split the boxes up based on `major_category`. Looking at the box plot proved my hypothesis correct, as the boxes for Transportation, Sales, and Finances were noticeably lower than the other boxes. Furthermore, computing jobs had a much higher 25th percentile compared to the others. Combined with the large variety of whiskers on the plots, I felt that `major_category` was a good variable to include in my model.

```{r major_cat, echo=FALSE, warning=FALSE}
ggplot(data = training_data, mapping=aes(x=wage_percent_of_male, fill=major_category)) +
  geom_boxplot() +
  labs(title="Wage Percentage Compared to Category",
       x="Wage percent of male",
       fill="Category")+
  theme(
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()
  )+
  scale_fill_viridis_d()
```

### Building my Model

Combining all of my variables into a single linear model gets the equation:

wage_percent_of_male-hat = -828 + (0.046 * percent_female) + (0.452 * year) + (2.09 * Computing) + (1.80 * Construction) + (-4.71 * Finances) + (1.88 * Healthcare) + (-2.20 * Sales) + (1.46 * Service) + (-3.39 * Transportation)

```{r Lorenzo_model, echo=FALSE}
Lorenzo_model <- lm(wage_percent_of_male ~ percent_female + year + major_category, data = training_data)

Lorenzo_model %>% tidy()
```
```{r Lorenzo_stats, include=FALSE}
glance(Lorenzo_model) %>% select(r.squared, adj.r.squared)

Lorenzo_pred <- predict(Lorenzo_model, newdata=test_data) %>%
  bind_cols(test_data %>% select(wage_percent_of_male)) %>%
  rename(pred = ...1)

rmse(Lorenzo_pred, truth=wage_percent_of_male, estimate=pred)
```

## Model proposed by Sonia

## Model proposed by Jingrui

## Models Comparison

Creator of Model  | Adjusted R-squared | RMSE
------------- | ------------- | --------
Lorenzo  | 0.0689  |  8.6708
Sonia | 0.0724 | 8.6483
Jingrui | 0.5997 | 6.0623

comparison, higher r-squared is better, lower RMSE is better.

# Results

```{r full_data, inlcude=FALSE}
best_pred <- predict(best_model, newdata=jobs) %>%
  bind_cols(jobs %>% select(wage_percent_of_male)) %>%
  rename(pred = ...1)

rmse(best_pred, truth=wage_percent_of_male, estimate=pred)
```

...

